### 2022-07-27 paper/figures/IRV2DRPlot.tex

\begin{figure}[h]
    \begin{tikzpicture}
      \begin{axis}[xtick={0, 0.0003125, 0.000625, 0.00125, 0.0025, 0.005, 0.01}, x tick label style={rotate=45, log ticks with fixed point},xmode=log, log basis x=2, xlabel=Perturbation Radius ($\epsilon$), ylabel=Accuracy, width=\linewidth, height=7cm,legend style={at={(0.95,0.95)},anchor=north east}]

      \addplot[color=red,mark=x] coordinates {
        (0.0003125, 0.618)
        (0.000625, 0.286)
        (0.00125, 0.053)
        (0.0025, 0.007)
        (0.005, 0.0)
        (0.01, 0.0)
      };

      \addplot[color=blue,mark=*] coordinates {
        (0.0003125, 0.590)
        (0.000625, 0.255)
        (0.00125, 0.046)
        (0.0025, 0.0)
        (0.005, 0.0)
        (0.01, 0.0)
      };

      \addplot[color=red, domain=0.0003125:0.01, dashed]{0.889};
      \addplot[color=blue, domain=0.0003125:0.01, dashed]{0.887};

      \legend{Without Attention,With Attention}
      \end{axis}
      \end{tikzpicture}
    \caption{Clean- and robust-accuracy of InceptionResNetV2 models for diabetic retinopathy detection. Dashed lines represent clean-accuracy.}
    \label{IRV2DRRobustness}
  \end{figure}