### 2022-07-27 paper/figures/IRV2PneumoPlot.tex

\begin{figure}[h]
    \begin{tikzpicture}
      \begin{axis}[xtick={0, 0.0003125, 0.000625, 0.00125, 0.0025, 0.005, 0.01}, x tick label style={rotate=45, log ticks with fixed point},xmode=log, log basis x=2, xlabel=Perturbation Radius ($\epsilon$), ylabel=Accuracy, width=\linewidth, height=7cm,legend style={at={(0.95,0.95)},anchor=north east}]

      \addplot[color=red,mark=x] coordinates {
        (0.0003125, 0.984)
        (0.000625, 0.982)
        (0.00125, 0.969)
        (0.0025, 0.945)
        (0.005, 0.740)
        (0.01, 0.137)
      };

      \addplot[color=blue,mark=*] coordinates {
        (0.0003125, 0.972)
        (0.000625, 0.949)
        (0.00125, 0.817)
        (0.0025, 0.334)
        (0.005, 0.334)
        (0.01, 0.0)
      };

      \addplot[color=red, domain=0.0003125:0.01, dashed]{0.986};
      \addplot[color=blue, domain=0.0003125:0.01, dashed]{0.987};

      \legend{Without Attention,With Attention}
      \end{axis}
      \end{tikzpicture}
    \caption{Clean- and robust-accuracy of InceptionResNetV2 models for pneumothorax detection. Dashed lines represent clean-accuracy.}
    \label{IRV2PneumoRobustness}
  \end{figure}